---
title: "Generate Metabolite library entries for MetaboAnnotatoR package"
author: "Gonçalo Graça"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generate Metabolite library entries for MetaboAnnotatoR package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r options, include=FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

## This vignette illustrates how to generate a metabolite library entry from 
an experimental MS/MS spectrum.

First load library and dependencies:

```{r load_package, eval=TRUE, message=FALSE}
library(MetaboAnnotatoR)
```

### Read the MS/MS spectrum

First we set the working directory (to tempdir() in this example) and then
read the spectrum as data frame into R. In this example we will use the MS/MS
spectrum of D-Pantothenic Acid [M+H]+ adduct from MassBank, accession code:
[MSBNK-RIKEN-PR100295](https://massbank.eu/MassBank/RecordDisplay?id=MSBNK-RIKEN-PR100295):

```{r import_spectrum, eval=TRUE}
setwd(tempdir())
specObject <- data.frame(V1=c(70.0298, 85.0652, 90.0556, 98.024, 116.0353,
                            124.0766, 184.0981, 202.1085, 220.1185),
                        V2=c(13.965907, 13.534607, 100.0, 26.165537, 
                            15.383036, 25.231054, 28.578764, 43.017047, 
                            64.962005))
```


### Generate library entry

We can then convert the spectra into a library entry by using the function 
'genFragEntry'. This function will attribute occurrence scores to the peaks
above 'mpeaksThres' threshold ("marker peaks") and above the noise level.
Note that metabolite name, adduct name, accurate adduct m/z and output filename
need to be explicitly defined.
The peak-picking parameters 'noise' and 'mpeaksThres' and the occurrence scores
used here are the default parameters.

```{r run_function, eval=TRUE}
genFragEntry(specObject, 
            "Pantothenic acid",
            "[M+H]+",
            220.1179,
            "Pantothenic_acid_pos.csv",
            noise=0.005,
            mpeaksScore=0.9, 
            mpeaksThres=0.1,
            mzTol=0.01)
```

A .csv file named "Pantothenic_acid_pos.csv" will be generated in the working 
directory. This should be added to the local "Libraries" folder, under the 
path "/Libraries/Metabolites/POS".

# Session Information
```{r session_info, , eval=TRUE, echo=FALSE}
devtools::session_info()
```