---
title: "Generate Metabolite library entries for MetaboAnnotatoR package"
author: "Gonçalo Graça"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{generate library entry}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## This vignette illustrates how to generate a metabolite library entry from an experiemntal MS/MS spectrum

First load library and dependencies:

```{r, eval = FALSE}
library(MetaboAnnotatoR)
```

### Read the MS/MS spectrum

First we read the spectrum saved as .txt into R. In this example we will use the MS/MS spectrum of D-Pantothenic Acid [M+H]+ adduct
from MassBank (accession code: PR100295). This file has been included in as example data folder of MetaboAnnotatoR:

```{r, eval = FALSE}
spec <- system.file("/Data/Pantothenic_acid_pos.txt", package = "MetaboAnnotatoR")
specObject <- read.table(spec, header=FALSE)
```


### Generate library entry

We can then convert the spectra into a library entry by using function genFragEntry.
This function will attribute occurrence scores to the peaks above 'mpeaksThres' threshold ("marker peaks") and above the noise level.
Note that metabolite name, adduct name, accurate adduct m/z and output filename need to be explicitly defined.
The peak-picking parameters 'noise' and 'mpeaksThres' and the occurrence scores used here are the default parameters.

```{r, eval = FALSE}
genFragEntry(specObject, 
             "Pantothenic acid",
             "[M+H]+",
             220.1179,
             "Pantothenic_acid_pos.csv",
             noise = 0.005,
             mpeaksScore = 0.9, 
             mpeaksThres = 0.1,
             mzTol = 0.01)
```

A .csv file named "Pantothenic_acid_pos.csv" will be generated in the working directory. This should be added to the local "Libraries" folder, under the path /Libraries/Metabolites/POS.