---
title: "Import an MS/MS library"
author: "Gonçalo Graça"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Import an MS/MS library}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r options, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

# Introduction
This vignette illustrates how to convert MS/MS spectra libraries 
(.msp files) to MetaboAnnotatoR metabolite library entries.

# Import .msp library

**Load MetaboAnnotatoR library and dependencies**

```{r load_package, eval=TRUE, message=FALSE, warning=FALSE}
library(MetaboAnnotatoR)
```

**Get the path to the example .msp MS/MS spectra library file:**

Get the path to the example file containing the .msp library. 
For illustrative purposes, an .msp file containing some MS/MS spectra obtained
from MassBank from both positive and negative ionisation modes was used:

```{r filepath, eval=TRUE}
msp_path <- system.file("MassBank_example.msp", 
                        package="MetaboAnnotatoR")
```

**Set a directory to store


**Generate library entries imported from the .msp file**

For this example a temporary directory will be used, but the user is encouraged
to use any other preferred directory.
```{r setwd_to_run, eval=TRUE}
LibDir <- tempdir()
```

Use function *mspToLib* to read and convert the spectra records from the 
.msp files into library entries (.csv files), which will be stored in the 
user-defined library directory (*LibDir*). A "positive" or "negative" mode
suffix will be added to each file name to facilitate the organisation of the 
custom libraries.
Similarly to the function *genFragEntry* this function will attribute 
occurrence scores to the peaks above *mpeaksThres* threshold 
("marker peaks") and above the noise level. The peak-picking parameters 
*noise* and *mpeaksThres* and the occurrence scores used here are the
default parameters.

```{r run_function, eval=TRUE, warning=FALSE}
mspToLib(msp_path, LibDir, noise=0.005, mpeaksScore=0.9, mpeaksThres=0.1)
```

# Session Info

```{r sessionInfo}
sessionInfo()
```