---
title: "Introduction to MetaboAnnotatoR"
author: "Gonçalo Graça"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Introduction to MetaboAnnotatoR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r doc_options, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

# Introduction

MetaboAnnotatoR is designed to perform metabolite annotation of features from
LC-MS All-ion fragmentation (AIF) datasets, using ion fragment databases.
It requires raw LC-MS AIF chromatograms acquired/transformed in centroid mode.

# Installation

To install this package, start R (version "4.5") and enter:
```{r installation, eval=FALSE, echo=TRUE, message=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("MetaboAnnotatoR")
```


# Example session

An example of feature annotation using LC-MS AIF chromatograms processed 
using xcms and RamClustR packages is illustrated here.
The details of how the example dataset was obtained check MetaboAnnotatoR
original paper for the full details: 
https://pubs.acs.org/doi/10.1021/acs.analchem.1c03032.

For more details on RAMClustR object, check the original publication:
https://pubs.acs.org/doi/10.1021/ac501530d.  

Firstly load library and dependencies:

```{r load_package, eval=TRUE, echo=TRUE, message=FALSE}
library(MetaboAnnotatoR)
```

```{r setwd_to_run, eval=TRUE, echo=FALSE, message=FALSE}
## silently set working directory to tempdir() for the vignette
setwd(tempdir())
```

## Getting feature table and example data

As an input the software requires a .csv file containing the features to be 
annotated (targetTable.csv - the file name can be changed).
The following function will place the example 'targetTable.csv' and the 
'XCMS_options.csv' (used for data peak-picking, when dealing with raw data) 
files in the user working directory. The option AIFdata=TRUE will load the 
AIF processed data into the global environment. These consist of two objects: 
1) an xcmsSet object (XSet) containing the processed data from 100 AIF LC-MS 
chromatograms from human serum samples and 
2) the respective pseudo-MS/Ms spectra obtained by processing the xcmsSet 
data using RAMClustR (RC): 

```{r load_data, eval=TRUE, results='hide'}
getDemoData(AIFdata=TRUE)
```


The 'targetTable.csv' file contains 6 features, corresponding to 
4 phospholipids and 2 lipid-like molecules. 
Annotation can be performed using two of the default libraries: 'Lipids'
or 'Metabolites' of either ESI+ (POS) or ESI- (NEG) ionisation modes.

## Annotations using the Lipid libraries

Let's start the annotation by searching the Lipid libraries:

```{r lipids_example, eval=TRUE}
annotateRC(targetTable="targetTable.csv", 
            xcmsObject=Xset, 
            ramclustObj=RC, 
            libs="Lipids", 
            ESImode="POS")
```

A new folder will be automatically created in the working directory called
'Annotations' were annotation results will be stored. Several files will be
present in this folder. The annotation results are stored in .csv files for 
each feature. The annotated 'targetTable' can be found in the 
'Global_Results.csv' file. Graphical results are stored as .pdf files in 
the same folder.

Some features will remain un-annotated, which means that no matching compound
was found in the lipids library. 

## Annotations using the non-lipid Metabolites libraries

We can now try to use the non-lipid molecule libraries to annotate :

```{r metabolites_example, eval=TRUE}
annotateRC(targetTable="targetTable.csv", 
            xcmsObject=Xset, 
            ramclustObj=RC, 
            libs="Metabolites", 
            ESImode="POS")
```

The libraries used in the example are the default 'Lipids' and 'Metabolites' 
(ESI+ mode) libraries provided with the package.
Similarly to the lipid annotations, an annotations folder will be created in 
the working directory to store the annotations obtained using the 'Metabolites'
ESI+ library.

# Session Information

```{r session_info, , eval=TRUE, echo=FALSE}
devtools::session_info()
```